EDA

### 2020-03-01 src/EDA&mode_fitting.Rmd
models using orginal value are updated now.

model fitting
(i) linear model
Not too much difference between linear 1 and linear 2 (with and without year).
This probably implies adding year doesn't make a difference.
```{r}
linear1<-lm(rate_pct~factor(Year)+factor(TaxClassCode)+factor(Municipalities)+assessTotal_pct+tax,data=pct_final)
summary(linear1)

# linear1<-lm(rate~factor(Year)+factor(TaxClassCode)+factor(AddressAssessorMunicipalityDesc)+assessTotal+predTaxIncome,data=assessment_aggregate)
# summary(linear1)

linear<-lm(rate~.,data=assessment_aggregate)
summary(linear)

linear2<-lm(rate_pct~factor(TaxClassCode)+factor(Municipalities)+assessTotal_pct+tax,data=pct_final)
summary(linear2)

# goodness of fit (all data)
library(broom)
linear1_fit<-augment(linear1)
sqrt(sum((linear1_fit$.resid)^2)/nrow(pct_final)) # 6.672786
linear2_fit<-augment(linear2)
sqrt(sum((linear2_fit$.resid)^2)/nrow(pct_final)) # 7.248359

```
linear 1: Residual standard error: 7.296 on 138 degrees of freedom
Multiple R-squared:  0.265,	Adjusted R-squared:  0.1265 
PMSE = 6.672786

linear 2: Residual standard error: 7.869 on 140 degrees of freedom
Multiple R-squared:  0.1327,	Adjusted R-squared:  -0.01593 
F-statistic: 0.8929 on 24 and 140 DF,  p-value: 0.6112
PMSE = 7.248359

we re-looked at the data and found percentage changes of tax in "Langley - City" and "Langley - Township" are too large (~400). We ignored the two municipalities and refit the linear model

```{r}
test<-pct_final %>% filter(Municipalities!="Langley - City") %>% 
  filter(Municipalities!="Langley - Township")
linear_test<-lm(rate_pct~factor(Year)+factor(TaxClassCode)+factor(Municipalities)+assessTotal_pct+tax,data=test)
summary(linear_test)

# goodness of fit
fit_linear_test<-augment(linear_test)
sqrt(sum((fit_linear_test$.resid)^2)/nrow(test)) # 6.767493

linear_test_2<-lm(rate_pct~factor(TaxClassCode)+factor(Municipalities)+assessTotal_pct+tax,data=test)
summary(linear_test_2)

# goodness of fit
fit_linear_test_2<-augment(linear_test_2)
sqrt(sum((fit_linear_test_2$.resid)^2)/nrow(test)) #  7.50047

```
After filtering Langely
with variable year:
Residual standard error: 7.429 on 122 degrees of freedom
Multiple R-squared:  0.2974,	Adjusted R-squared:  0.1592 
F-statistic: 2.152 on 24 and 122 DF,  p-value: 0.003567
PMSE: 6.767493

without variable year
Residual standard error: 8.167 on 124 degrees of freedom
Multiple R-squared:  0.137,	Adjusted R-squared:  -0.01609 
F-statistic: 0.8949 on 22 and 124 DF,  p-value: 0.6014
PMSE: 7.50047

```{r}
linear3<-lm(rate_pct~factor(Year)+factor(TaxClassCode)+factor(Municipalities),data=pct_final)
summary(linear3)
fit_linear3<-augment(linear3)
sqrt(sum((fit_linear3$.resid)^2)/nrow(test)) #  7.50047

Question: How to judge goodness of fit? Multiple R square? Adjusted R square? or mean prediction error?
**TODO: create training/test sets to test the prediction power!

@@

Ridge Regression
reference: https://www.pluralsight.com/guides/linear-lasso-and-ridge-regression-with-r
```{r}
library(glmnet)
library(dummies)
dummy_year<-dummy(test$Year)
dummy_municipal<-dummy(test$Municipalities)
dummy_taxclass<-dummy(test$TaxClassCode)
# build x matrix
x<-cbind(dummy_year,dummy_municipal)
x<-cbind(x,dummy_taxclass)
x<-cbind(x,as.numeric(test$assessTotal_pct))
x<-cbind(x,as.numeric(test$tax))
y<-test$rate_pct
lambdas <- 10^seq(2, -3, by = -.1)
dim(x)
colnames(x)[26]<-paste("assessment")
colnames(x)[27]<-paste("tax")
colnames(x)

@@

# Compute R^2 from true and predicted values
eval_results <- function(true, predicted) {
  SSE <- sum((predicted - true)^2)
  SST <- sum((true - mean(true))^2)
  R_square <- 1 - SSE / SST
  RMSE = sqrt(SSE/nrow(predicted))
# Model performance metrics
data.frame(
  RMSE = RMSE,
  Rsquare = R_square
)

@@

```
RMSE=7.159649
Rsquare=0.2136659
Ridge regression doesn't seem to work well.

Lasso 
```{r}
lambdas <- 10^seq(2, -3, by = -.1)

# Setting alpha = 1 implements lasso regression
set.seed(450)
lasso_reg <- cv.glmnet(x, y, alpha = 1, lambda = lambdas, standardize = TRUE, nfolds = 10)

@@

predictions_train <- predict(lasso_model, s = lambda_best, newx = x)
eval_results(y, predictions_train)
```
RMSE = 7.221299, Rsquare=0.2000649

Elastic net: 
reference: https://daviddalpiaz.github.io/r4sl/elastic-net.html
```{r}
library(caret)
tibble::as_tibble(test)
cv_10 = trainControl(method = "cv", number = 10)
ridge = train(
  rate_pct ~ Year + Municipalities + TaxClassCode + assessTotal_pct + tax, data = test,
  method = "glmnet",
  trControl = cv_10
)
ridge
```
apha = 1 which implies lasso! refer to lasso.

Prediction power
train/test
```{r}
set.seed(450)
train_ind<-sample(165,120)
train1<-pct_final[train_ind,]
test1<-pct_final[-train_ind,]
```

linear model
```{r}
# full model with year
newx<-test1[,-c(5,6,7,8)]
y<-test1[,c(8)]
linear_1<-lm(rate_pct~factor(Year)+factor(TaxClassCode)+factor(Municipalities)+assessTotal_pct+tax,data=train1)
resid<-predict(linear_1,newdata = newx) - y
sqrt(sum(resid^2)/nrow(test1))
# pmse = 6.5688

# full model without year
linear_2<-lm(rate_pct~factor(TaxClassCode)+factor(Municipalities)+assessTotal_pct+tax,data=train1)
resid<-predict(linear_2,newdata = newx) - y
sqrt(sum(resid^2)/nrow(test1))
# 6.855429

# reduced variable
linear_3<-lm(rate_pct~factor(Year)+factor(TaxClassCode)+factor(Municipalities),data=train1)
resid<-predict(linear_3,newdata = newx) - y
sqrt(sum(resid^2)/nrow(test1))
# 6.570848
```
Lasso
```{r}
# create the whole matrix
dummy_year<-dummy(pct_final$Year)
dummy_municipal<-dummy(pct_final$Municipalities)
dummy_taxclass<-dummy(pct_final$TaxClassCode)
# build x matrix
x<-cbind(dummy_year,dummy_municipal)
x<-cbind(x,dummy_taxclass)
x<-cbind(x,as.numeric(pct_final$assessTotal_pct))
x<-cbind(x,as.numeric(pct_final$tax))
y<-as.matrix(pct_final$rate_pct)
dim(x) # 165  29
dim(y)
# creat x_train matrix and y_train
x_train<-x[train_ind,]
y_train<-y[train_ind,]
# create x_test matrix
x_test<-x[-train_ind,]
y_test<-y[-train_ind,]

lambdas <- 10^seq(2, -3, by = -.1)
# Setting alpha = 1 implements lasso regression
set.seed(450)
lasso_reg <- cv.glmnet(x_train, y_train, alpha = 1, lambda = lambdas, standardize = TRUE, nfolds = 10)

@@

predictions_test <- predict(lasso_model, s = lambda_best, newx = x_test)
eval_results(y_test, predictions_test)
# 7.012128	
```

Ridge
```{r}
ridge_reg = glmnet(x_train, y_train, nlambda = 25, alpha = 0, family = 'gaussian', lambda = lambdas)
set.seed(450)
cv_ridge <- cv.glmnet(x_train, y_train, alpha = 0, lambda = lambdas, nfolds=10)
optimal_lambda <- cv_ridge$lambda.min
optimal_lambda
predictions_test <- predict(ridge_reg, s = optimal_lambda, newx = x_test)
eval_results(y_test, predictions_test)
# 6.885151	
```

>>>>

model fitting
(i) linear model
```{r}
linear1<-lm(rate_pct~factor(Year)+factor(TaxClassCode)+factor(Municipalities)+assessTotal_pct+tax,data=pct_final)
summary(linear1)

linear_full<-lm(rate~.,data=assessment_aggregate)
summary(linear_full)

library(broom)
linear_full_fit<-augment(linear_full)
sqrt(sum((linear_full_fit$.resid)^2)/nrow(assessment_aggregate)) #  2.772333

```
Multiple R-squared:  0.7803,	Adjusted R-squared:  0.7503 
MSPE = 2.772333

Next: variable selection
with only significant variables
```{r}
reduced<-lm(rate~Year+TaxClassCode+AddressAssessorMunicipalityDesc, data=assessment_aggregate)
summary(reduced)

reduced_fit<-augment(reduced)
sqrt(sum((reduced_fit$.resid)^2)/nrow(assessment_aggregate)) #  2.861619

```
Multiple R-squared:  0.7659,	Adjusted R-squared:  0.7408 
MSPE = 2.861619
With fewer variables, adjusted R-squared declines.

By comparing the two models, the fit of full model is better.

@@

Ridge Regression
reference: https://www.pluralsight.com/guides/linear-lasso-and-ridge-regression-with-r
```{r}
library(glmnet)
library(dummies)
dummy_year<-dummy(assessment_aggregate$Year)
dummy_municipal<-dummy(assessment_aggregate$AddressAssessorMunicipalityDesc)
dummy_taxclass<-dummy(assessment_aggregate$TaxClassCode)
# build x matrix
x<-cbind(dummy_municipal,dummy_year,dummy_taxclass,assessment_aggregate$assessTotal,assessment_aggregate$landTotal,assessment_aggregate$improvementTotal,assessment_aggregate$propertyCount,assessment_aggregate$predTaxIncome,assessment_aggregate$tax)

y<-assessment_aggregate$rate
lambdas <- 10^seq(2, -3, by = -.1)
dim(x)

@@

# Compute R^2 from true and predicted values
eval_results <- function(true, predicted) {
  SSE <- sum((predicted - true)^2)
  SST <- sum((true - mean(true))^2)
  R_square <- 1 - SSE / SST
  MSPE = sqrt(SSE/nrow(predicted))
# Model performance metrics
data.frame(
  MSPE = MSPE,
  Rsquare = R_square
)

@@

```
MSPE = 1.992244
Rsquare=0.8865213

Lasso 
```{r}
# Setting alpha = 1 implements lasso regression
set.seed(450)
lasso_reg <- cv.glmnet(x, y, alpha = 1, lambda = lambdas, standardize = TRUE, nfolds = 10)

@@

predictions_train <- predict(lasso_model, s = lambda_best, newx = x)
eval_results(y, predictions_train)
```
MSPE = 1.985639
Rsquare=0.8872724

Elastic net: 
reference: https://daviddalpiaz.github.io/r4sl/elastic-net.html
```{r}
library(caret)
tibble::as_tibble(assessment_aggregate)
cv_10 = trainControl(method = "cv", number = 10)
elastic_net = train(
  rate~., data = assessment_aggregate,
  method = "glmnet",
  trControl = cv_10
)
elastic_net
```
The final values used for the model were alpha = 0.1 and lambda = 0.09334029.
MSPE = 3.066083  
R squared = 0.7459584

Among these all the models, lasso seems to have the best goodness of fit.

Prediction power
train/test
```{r}
set.seed(450)
train_ind<-sample(218,218-50)
train<-assessment_aggregate[train_ind,]
test<-assessment_aggregate[-train_ind,]
```

Full linear model
```{r}

newx<-test[,-c(8)]
y<-test[,c(8)]
linear_1<-lm(rate ~.,data=train)
resid<-predict(linear_1,newdata = newx) - y
sqrt(sum(resid^2)/nrow(test))
```
MSPE = 3.52065

Reduced model
```{r}
# reduced variable
linear_2<-lm(rate~Year+TaxClassCode+AddressAssessorMunicipalityDesc,data=train)
resid<-predict(linear_2,newdata = newx) - y
sqrt(sum(resid^2)/nrow(test))
# 3.516749
```
MSPE = 3.516749

Lasso
```{r}
# create the whole matrix
y<-as.matrix(assessment_aggregate$rate)
dim(x) # 165  29
dim(y)
# creat x_train matrix and y_train
x_train<-x[train_ind,]
y_train<-y[train_ind,]
# create x_test matrix
x_test<-x[-train_ind,]
y_test<-y[-train_ind,]

# Setting alpha = 1 implements lasso regression
set.seed(450)
lasso_reg <- cv.glmnet(x_train, y_train, alpha = 1, lambda = lambdas, standardize = TRUE, nfolds = 10)

@@

predictions_test <- predict(lasso_model, s = lambda_best, newx = x_test)
eval_results(y_test, predictions_test)
```
MSPE = 2.530686

Ridge
```{r}
ridge_reg = glmnet(x_train, y_train, nlambda = 25, alpha = 0, family = 'gaussian', lambda = lambdas)
set.seed(450)
cv_ridge <- cv.glmnet(x_train, y_train, alpha = 0, lambda = lambdas, nfolds=10)
optimal_lambda <- cv_ridge$lambda.min
optimal_lambda
predictions_test <- predict(ridge_reg, s = optimal_lambda, newx = x_test)
eval_results(y_test, predictions_test)
```
MSPE = 2.568047

Elastic Net
```{r}
tibble::as_tibble(assessment_aggregate[train_ind,])
cv_10 = trainControl(method = "cv", number = 10)
elastic_net = train(
  rate~., data = assessment_aggregate[train_ind,],
  method = "glmnet",
  trControl = cv_10
)
elastic_net
# RMSE was used to select the optimal model using the smallest value.
# The final values used for the model were alpha = 0.55 and lambda = 0.009140377.

ridge_reg = glmnet(x_train, y_train, nlambda = 25, alpha = 0.55, family = 'gaussian', lambda = 0.009140377)
predictions_test <- predict(ridge_reg, newx = x_test)
eval_results(y_test, predictions_test)
```
MSPE = 2.566096
